Public transport and school location impacts on educational inequalities: insights from São Paulo

### 2017-04-02 paper_v2.tex
Add new reference
Veitch, J., A. Carver, J. Salmon, G. Abbott, K. Ball, D. Crawford, V. Cleland, and A. Timperio. 2017. “What Predicts Children’s Active Transport and Independent Mobility in Disadvantaged Neighborhoods?” Health & Place 44 (March): 103–9. doi:10.1016/j.healthplace.2017.02.003.

\begin{abstract}
In many large Latin American urban areas such as the São Paulo
Metropolitan Region (SPMR), growing social and economic inequalities are embedded through high spatial inequality in the provision of state schools and affordable public transport to these schools. This paper sheds light on the transport-education inequality nexus with reference to school accessibility in the SPMR. 
To assess school accessibility, we develop an accessibility index which combines information on the spatial distribution of adolescents, the location of existing schools, and the public transport provision serving the school catchment area into a single measure. The index is used to measure school accessibility locally across 633 areas within the SPMR. We use the index to simulate the impact of a policy aiming at increasing the centralization of public secondary education provision, and find that it negatively affects students with the lowest levels of accessibility. These results illustrate how existing inequalities can be amplified by variable accessibility to schools across income groups and geographical space.
The research suggests that educational inequality impacts of school agglomeration policies should be considered before centralisation takes place and highlights the links between transport planning social injustice that.
\end{abstract}

@@

The SPMR presents an excellent opportunity to measure the extent of
educational and transport infrastructure, and to study the impact of
public policies on different socio-economic groups. Partly as a response to the pressures from the Free Pass Movement that started in March, 2013, low income public school students residing in the SPMR gained access to a public transport subsidy in 2015. However, in October 2015, the São Paulo state government announced during a television interview that as part of a budgetary deficit reduction plan, dozens of secondary schools would be closed in 2016. The change in policy was estimated to affect more than 300,000 students, many whom would be placed in schools far from their homes. As a response, in the following months students occupied over 200 schools. Protesters scored some victories: the reform was postponed for one year, and the education secretary resigned (Ortellado 2015).

@@

>>>>

\begin{abstract}
In many large Latin American urban areas such as the São Paulo
Metropolitan Region (SPMR), growing social and economic inequalities are embedded through high spatial inequality in the provision of state schools and affordable public transport to these schools. This paper sheds light on the transport-education inequality nexus with reference to school accessibility by public transport in the SPMR. 
To assess school accessibility, we develop an accessibility index which combines information on the spatial distribution of adolescents, the location of existing schools, and the public transport provision serving the school catchment area into a single measure. The index is used to measure school accessibility locally across 633 areas within the SPMR. We use the index to simulate the impact of a policy aiming at increasing the centralization of public secondary education provision, and find that it negatively affects public transport accessibility for students with the lowest levels of accessibility. These results illustrate how existing inequalities can be amplified by variable accessibility to schools across income groups and geographical space.
The research suggests that educational inequality impacts of school agglomeration policies should be considered before centralisation takes place and highlights the links between transport planning social injustice that.
\end{abstract}

@@

The SPMR presents an excellent opportunity to measure the extent of
educational and public transport provision, and to study the impact of
public policies on different socio-economic groups. Partly as a response to the pressures from the Free Pass Movement that started in March, 2013, low income public school students residing in the SPMR gained access to a public transport subsidy in 2015. However, in October 2015, the São Paulo state government announced during a television interview that as part of a budgetary deficit reduction plan, dozens of secondary schools would be closed in 2016. The change in policy was estimated to affect more than 300,000 students, many whom would be placed in schools far from their homes. As a response, in the following months students occupied over 200 schools. Protesters scored some victories: the reform was postponed for one year, and the education secretary resigned (Ortellado 2015).
It was this policy context that informed the paper's focus on public transport, rather than the topic of active travel to school, which has gained recent prominence in Australia (e.g. Veitch et al., 2017) the USA and Europe

@@

Veitch, J., A. Carver, J. Salmon, G. Abbott, K. Ball, D. Crawford, V. Cleland, and A. Timperio. 2017. ``What Predicts Children’s Active Transport and Independent Mobility in Disadvantaged Neighborhoods?'' \emph{Health \& Place} 44 (March): 103–9. doi:10.1016/j.healthplace.2017.02.003.
